---
author: "Mehrdad Arab"
date: "November 30, 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=7, fig.height=4, warning=FALSE, message=FALSE)
```

# Statistical Inference Course Project (Part_1)

## Overview
This project will investigate the exponential distribution in R and compare it with the Central Limit Theorem. Given that lambda = 0.2 for all of the simulations. Part 1 of the project will investigate the distribution of averages of 40 exponentials over a thousand simulations.

## Simulations

```{r}
# load neccesary libraries
library(ggplot2)
# Using pre-defined parameters
lambda <- 0.2
n <- 40
sims <- 1:1000
set.seed(123)

# Simulate the population
population <- data.frame(x=sapply(sims, function(x) {mean(rexp(n, lambda))}))

# Plot the histogram
hist.pop <- ggplot(population, aes(x=x)) + 
  geom_histogram(aes(y=..count.., fill=..count..)) +
  labs(title="Histogram for Averages of 40 Exponentials over 1000 Simulations", y="Frequency", x="Mean")
hist.pop
```


## Sample Mean versus Theoretical Mean

The expected mean $\mu$ of a exponential distribution of rate $\lambda$ is 

$\mu= \frac{1}{\lambda}$ 

```{r}
theoretical.mean <- 1/lambda
theoretical.mean
```

Let $\bar X$ be the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.

```{r}
# Tabulating the Sample Mean & Theoretical Mean
sample.mean <- mean(population$x)
cbind(sample.mean, theoretical.mean)
```

As you can see the expected mean and the avarage sample mean are very close 


## Sample Variance versus Theoretical Variance



```{r}
sample.variance <- var(population$x)
theoretical.variance <- ((1/lambda)^2)/n
cbind(sample.variance, theoretical.variance)
```

As we can see below both Sample Variance and Theoretical Variance are very close.


## Distribution


```{r echo=FALSE}
# Plotting Sample Mean & Varience vs Theoretical Mean & Varience
gg <- ggplot(population, aes(x=x)) +
  geom_histogram(aes(y=..density.., fill=..density..)) +
  labs(title="Histogram of Averages of 40 Exponentials over 1000 Simulations", y="Density", x="Mean") + 
  geom_density(colour="blue") +
  geom_vline(xintercept=sample.mean, colour="blue", linetype="dashed") +
  stat_function(fun=dnorm,args=list( mean=1/lambda, sd=sqrt(theoretical.variance)),color = "red") +
  geom_vline(xintercept=theoretical.mean, colour="red", linetype="dashed") 
gg
```

As we can see, the Sampled mean for 40 exponentials simulated 1000 times are very close to the Theoretical mean for a normal distribution.

